---
title: "Dissertation Analysis"
author: "CM793"
date: "2023-04-23"
always_allow_html: true
output:
  pdf_document: default
  html_document: default
  
---

1. The predicted values from the model (x-axis) against the residuals (y-axis). Use     this plot to look for linearity and homoscedasticity.
2. A Q-Q plot of the standardized residuals. Use this plot to look for normality of    residuals.
3. The predicted values from the model (x-axis) against the square root of the     standardized residuals (y-axis). This is a variant of plot 1 and is used to look  for linearity and homoscedasticity.
4. The case number (x-axis) against the Cooks distance (y-axis). This plot can help    to identify influential cases (cases with large values for Cooks distance).
5. The leverage value for each case (x-axis) against the standardized residual  (y-axis). This plot is used to identify influential cases and outliers. Leverage  values indicate the influence of an individual case on the model and are related to  Cook’s distance.
6. The leverage value for each case (x-axis) against the corresponding Cooks distance (y-axis). This plot is used to identify influential cases and outliers.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#Packages and data
library(knitr)
library(dplyr)
library(readxl)
library(readr)
library(tidyr)
library(plyr)
library(lavaan)
library(lavaanPlot)
library(ltm)
library(rmarkdown)
library(apaTables)
library(BayesFactor)
library(parameters)

#Read Data#
DATA_Xl <- readxl::read_excel('/cloud/project/data/Student perception of employability in elite professions_March 13, 2023_09.58.xlsx')
```

```{r}
#Clean Data

#Filter by completed responses
DATA_nona <- tidyr::drop_na(DATA_Xl, "Elite_Employ_16", "Prosociality & Polit_30")

#Filter by provided consent
DATA_con <- dplyr::filter(DATA_nona, C1==2, C2==2, C3==2, C4==2, C5==2, C6==2, C7==2)
```

```{r}
#Participant information viewing

#Subject
table(DATA_con$Subjects)

#Age by gender 1=Male 2=Female 3=Non-Binary
Num_info <- DATA_con %>% mutate_if(is.character, as.numeric)
Num_info %>% 
  dplyr::group_by(Gender) %>% 
  dplyr::summarise(
    Mean_age = mean(Age),
    SD_age = sd(Age),
    Number = dplyr::n()
  )

#Age
Num_info %>%
  dplyr::summarise(
    Mean_age = mean(Age),
    SD_age = sd(Age),
    Number = dplyr::n()
  )
```

```{r}
#Internal consistency of Pol/Soc and Employability

#Cronbachs alpha
SocScale <- data.frame(DATA_con[,c(102:107,111,113,118,120,122,124:125,129:130)])
ltm::cronbach.alpha(SocScale, na.rm=T)

PolScale <- data.frame(DATA_con[,c(108:110,112,114:117,119,121,123,126:128,131)])
ltm::cronbach.alpha(PolScale, na.rm=T)

EltScale <- data.frame(DATA_con[,c(172:187)])
ltm::cronbach.alpha(EltScale, na.rm=T)

#All three good internal consistency
```

```{r}
# Creating mean variables for Pol/Soc and Employability and tidy columns

#set as numeric
df <- DATA_con %>% mutate_if(is.character, as.numeric) 

#Pol & Soc means
df$Socmean <-apply(df[,c(102:107,111,113,118,120,122,124:125,129:130)],1,mean, na.rm=T, digits=3)
df$Polmean <-apply(df[,c(108:110,112,114:117,119,121,123,126:128,131)],1,mean, na.rm=T, digits=3)

#Elite employability means
df$Eltmean <-apply(df[,c(172:187)],1,mean, na.rm=T, digits=3)

#Standardised values -changes b but not p or f
df$Socmean <- (df$Socmean - mean(df$Socmean)) / sd(df$Socmean)
df$Polmean <- (df$Polmean - mean(df$Polmean)) / sd(df$Polmean)
df$Eltmean <- (df$Eltmean - mean(df$Eltmean)) / sd(df$Eltmean)

#Tidy columns
DATA_rel <- subset(df[ , c(18:28,132:144,154,188:190)])
                   
```

```{r}
#Evaluating and creating mean variables for SEC


#recode education and tuition
DATA_rel$Education <- recode(DATA_rel$Education, '2'=1, '3'=2, '1'=3)
DATA_rel$Tuition <- recode(DATA_rel$Tuition, '2'=1, '1'=2)

#standardised values
DATA_rel$StdEdu <- (DATA_rel$Education - mean(DATA_rel$Education)) / sd(DATA_rel$Education)
DATA_rel$StdTui <- (DATA_rel$Tuition - mean(DATA_rel$Tuition)) / sd(DATA_rel$Tuition)
DATA_rel$StdPar <- (DATA_rel$`Parents education` - mean(DATA_rel$`Parents education`)) / sd(DATA_rel$`Parents education`)
DATA_rel$StdInc <- (DATA_rel$`Household income` - mean(DATA_rel$`Household income`)) / sd(DATA_rel$`Household income`)
DATA_rel$StdCla <- (DATA_rel$`Social class` - mean(DATA_rel$`Social class`)) / sd(DATA_rel$`Social class`)
DATA_rel$StdLad <- (DATA_rel$`MacArthur Scale` - mean(DATA_rel$`MacArthur Scale`)) / sd(DATA_rel$`MacArthur Scale`)

#cronbach's alpha
ObjScale1 <- data.frame(DATA_rel[,c(29,30,31,32)])
ltm::cronbach.alpha(ObjScale1, na.rm=T) #Poor
ObjScale2 <- data.frame(DATA_rel[,c(30,31,32)])
ltm::cronbach.alpha(ObjScale2, na.rm=T) #good
ObjScale3 <- data.frame(DATA_rel[,c(31,32)])
ltm::cronbach.alpha(ObjScale3, na.rm=T) #worse
ObjScale4 <- data.frame(DATA_rel[,c(30,31)])
ltm::cronbach.alpha(ObjScale4, na.rm=T) #Best
#Tuition (30) and parents education (31) best internal consistency

SubScale <- data.frame(DATA_con[,c(33:34)])
ltm::cronbach.alpha(SubScale, na.rm=T)
#Poor internal consistency

#standardised coefficients 
Obj_lm1 <- lm(Eltmean ~ StdEdu + StdTui + StdPar + StdInc, data= DATA_rel, na.action= na.exclude) 
summary(Obj_lm1)
Obj_lm2 <- lm(Eltmean ~ StdTui + StdPar + StdInc, data= DATA_rel, na.action= na.exclude) #remove education
summary(Obj_lm2) #Mildly better R2, mid better P
Obj_lm3 <- lm(Eltmean ~ StdPar + StdInc, data= DATA_rel, na.action= na.exclude) #Remove parent education
summary(Obj_lm3) #Worse R2 and worse P
Obj_lm4 <- lm(Eltmean ~ StdTui + StdPar, data= DATA_rel, na.action= na.exclude) #Remove income
summary(Obj_lm4) #Best R2 and best P
#Concl = Tuition & parents education best

Sub_lm1 <- lm(Eltmean ~ StdLad, data= DATA_rel, na.action= na.exclude) 
summary(Sub_lm1) #wrse R2 and P
Sub_lm2 <- lm(Eltmean ~ StdCla, data= DATA_rel, na.action= na.exclude) 
summary(Sub_lm2)#better R2 and P
Sub_lm3 <- lm(Eltmean ~ StdLad*StdCla, data= DATA_rel, na.action= na.exclude) 
summary(Sub_lm3) #wrse R2 and P
#Concl = class better than ladder


#CONCLUSION: Only use tuition and parents education for objective, only use Class and ladder for subjective

#Creating aggregate value
DATA_rel$ObjClass <-apply(DATA_rel[,c(30:31)],1,mean, na.rm=T, digits=3) #Create Class aggregate value - standardised class values
DATA_rel$SubClass <-apply(DATA_rel[,c(33,34)],1,mean, na.rm=T, digits=3) #Create Class aggregate value - standardised class values

```

```{r}
#Assumption plotting

Asmp1 <- lm(Eltmean ~ ObjClass + Polmean + Socmean, data= DATA_rel, na.action= na.exclude)
plot(Asmp1, which = c(1,2,3,4,5,6)) #for obj
#Checking Normal residuals
summary(DATA_rel$Eltmean - Asmp1$fitted.values) #non linear by ~1/5b

Asmp2 <- lm(Eltmean ~ SubClass + Polmean + Socmean, data= DATA_rel, na.action= na.exclude)
plot(Asmp2, which = c(1,2,3,4,5,6)) #for sub
#Checking Normal residuals
summary(DATA_rel$Eltmean - Asmp2$fitted.values) #non linear by ~1/3b
```

```{r}
#Correlation matrix - Using unstandardised values

#set as numeric
apacor <- DATA_con %>% mutate_if(is.character, as.numeric) 

#Pol & Soc means
apacor$Socmean <-apply(apacor[,c(102:107,111,113,118,120,122,124:125,129:130)],1,mean, na.rm=T, digits=3)
apacor$Polmean <-apply(apacor[,c(108:110,112,114:117,119,121,123,126:128,131)],1,mean, na.rm=T, digits=3)
#Elite employability means
apacor$Eltmean <-apply(apacor[,c(172:187)],1,mean, na.rm=T, digits=3)

PearCorDat <- subset(apacor[ , c(23:28, 188:190)])


apacortab <- apa.cor.table(PearCorDat, filename = "CorTable.doc")


```

```{r}
#Simple 1 way ANOVAs
apaLmObj <- lm(Eltmean ~ ObjClass, data= DATA_rel, na.action= na.exclude)
apaLmSub <- lm(Eltmean ~ SubClass, data= DATA_rel, na.action= na.exclude)

parameters::model_parameters(apaLmObj, bootstrap = TRUE, digits = 3)

summary(apaLmObj)
summary(apaLmSub)
confint(apaLmObj)
confint(apaLmSub)
parameters::model_parameters(apaLmObj, bootstrap = TRUE, digits = 3)
parameters::model_parameters(apaLmSub, bootstrap = TRUE, digits = 3)


```


```{r}
#Mediation

#Mediation of polbeh & Eltemp for objective class
PolObjmodel <- "Eltmean ~ c*ObjClass + b*Polmean  
                   Polmean ~ a*ObjClass

                   indirect_effect := a*b
                   total_effect := c + (a*b)"

# fit the model with FIML and robust SEs
PolObjMed <- lavaan::sem(PolObjmodel, data = DATA_rel, missing = "FIML", estimator = "MLR")

#summarize the model (and round values for convenience)
broom::glance(PolObjMed) |>  knitr::kable(digits = 3)
broom::tidy(PolObjMed, conf.int = TRUE) |>  knitr::kable(digits = 3)
######################################################
######################################################
#Mediation of Socbeh & Eltemp for objective class
SocObjmodel <- "Eltmean ~ c*ObjClass + b*Socmean  
                   Socmean ~ a*ObjClass

                   indirect_effect := a*b
                   total_effect := c + (a*b)"

# fit the model with FIML and robust SEs
SocObjMed <- lavaan::sem(SocObjmodel, data = DATA_rel, missing = "FIML", estimator = "MLR")

#summarize the model (and round values for convenience)
broom::glance(SocObjMed) |>  knitr::kable(digits = 3)
broom::tidy(SocObjMed, conf.int = TRUE) |>  knitr::kable(digits = 3)
######################################################
######################################################
#Mediation of polbeh & Eltemp for subjective class
PolSubmodel <- "Eltmean ~ c*SubClass + b*Polmean  
                   Polmean ~ a*SubClass

                   indirect_effect := a*b
                   total_effect := c + (a*b)"

# fit the model with FIML and robust SEs
PolSubMed <- lavaan::sem(PolSubmodel, data = DATA_rel, missing = "FIML", estimator = "MLR")

#summarize the model (and round values for convenience)
broom::glance(PolSubMed) |>  knitr::kable(digits = 3)
broom::tidy(PolSubMed, conf.int = TRUE) |>  knitr::kable(digits = 3)
######################################################
######################################################
#Mediation of socbeh & Eltemp for subjective class
SocSubmodel <- "Eltmean ~ c*SubClass + b*Socmean  
                   Socmean ~ a*SubClass

                   indirect_effect := a*b
                   total_effect := c + (a*b)"

# fit the model with FIML and robust SEs
SocSubMed <- lavaan::sem(SocSubmodel, data = DATA_rel, missing = "FIML", estimator = "MLR")

#summarize the model (and round values for convenience)
broom::glance(SocSubMed) |>  knitr::kable(digits = 3)
broom::tidy(SocSubMed, conf.int = TRUE) |>  knitr::kable(digits = 3)
```

```{r}
#bootstrapped mediation 
######################################################
#Bootstrapped Mediation of polbeh & Eltemp for objective class
# fit the model with FIML and CHANGED TO NON ROBUST SE (ML)
PolObjBoot <- lavaan::sem(PolObjmodel, data = DATA_rel, test = "bootstrap",
                         bootstrap = 5000, estimator = "ML")

#summarize the model (and round values for convenience)
broom::glance(PolObjBoot) |>  knitr::kable(digits = 3)
broom::tidy(PolObjBoot, conf.int = TRUE) |>  knitr::kable(digits = 3)
######################################################
######################################################
#Bootstrapped Mediation of Socbeh & Eltemp for objective class
# fit the model with FIML and CHANGED TO NON ROBUST SE (ML)
SocObjBoot <- lavaan::sem(SocObjmodel, data = DATA_rel, test = "bootstrap",
                         bootstrap = 5000, estimator = "ML")

#summarize the model (and round values for convenience)
broom::glance(SocObjBoot) |>  knitr::kable(digits = 3)
broom::tidy(SocObjBoot, conf.int = TRUE) |>  knitr::kable(digits = 3)
######################################################
######################################################
#Bootstrapped Mediation of polbeh & Eltemp for subjective class
# fit the model with FIML and CHANGED TO NON ROBUST SE (ML)
PolSubBoot <- lavaan::sem(PolSubmodel, data = DATA_rel, test = "bootstrap",
                         bootstrap = 5000, estimator = "ML")

#summarize the model (and round values for convenience)
broom::glance(PolSubBoot) |>  knitr::kable(digits = 3)
broom::tidy(PolSubBoot, conf.int = TRUE) |>  knitr::kable(digits = 3)
######################################################
######################################################
#Bootstrapped Mediation of socbeh & Eltemp for subjective class
# fit the model with FIML and CHANGED TO NON ROBUST SE (ML)
SocSubBoot <- lavaan::sem(SocSubmodel, data = DATA_rel, test = "bootstrap",
                         bootstrap = 5000, estimator = "ML")

#summarize the model (and round values for convenience)
broom::glance(SocSubBoot) |>  knitr::kable(digits = 3)
broom::tidy(SocSubBoot, conf.int = TRUE) |>  knitr::kable(digits = 3)
```

```{r}
#Simple slopes analysis

library(ggeffects)
#For lower, median and upper quartile of behaviour
ss1 <- lm(Eltmean ~ ObjClass + Polmean + ObjClass:Polmean, data= DATA_rel, na.action= na.exclude)
ggpredict(ss1, terms = c("ObjClass[-2:2 by=0.1]", "Polmean[quart2]"), labels = labels1) |> plot(colors = "bw")

ss2 <- lm(Eltmean ~ ObjClass + Socmean + ObjClass:Socmean, data= DATA_rel, na.action= na.exclude)
ggpredict(ss2, terms = c("ObjClass[-2:2 by=0.1]", "Socmean[quart2]")) |> plot(colors = "bw")

ss3 <- lm(Eltmean ~ SubClass + Polmean + SubClass:Polmean, data= DATA_rel, na.action= na.exclude)
ggpredict(ss3, terms = c("SubClass[-2:2 by=0.1]", "Polmean[quart2]")) |> plot(colors = "bw")

ss4 <- lm(Eltmean ~ SubClass + Socmean + SubClass:Socmean, data= DATA_rel, na.action= na.exclude)
ggpredict(ss4, terms = c("SubClass[-2:2 by=0.1]", "Socmean[quart2]")) |> plot(colors = "bw")
```

```{r}
#Mediation Plotting

labels1 <- list(ObjClass = "Objective SEC", Polmean = "Political Behaviour", Eltmean = "Employability")
lavaanPlot(model = PolObjMed, labels = labels1, node_options = list(shape = "box", fontname = "Helvetica"), edge_options = list(color = "grey"), coefs = TRUE, digits=3, stars="regress")

labels2 <- list(ObjClass = "Objective SEC", Socmean = "Pro-social Behaviour", Eltmean = "Employability")
lavaanPlot(model = SocObjMed, labels = labels2, node_options = list(shape = "box", fontname = "Helvetica"), edge_options = list(color = "grey"), coefs = TRUE, digits=3, stars="regress")

labels3 <- list(SubClass = "Subjective SEC", Polmean = "Political Behaviour", Eltmean = "Employability")
lavaanPlot(model = PolSubMed, labels = labels3, node_options = list(shape = "box", fontname = "Helvetica"), edge_options = list(color = "grey"), coefs = TRUE, digits=3, stars="regress")

labels4 <- list(SubClass = "Subjective SEC", Socmean = "Pro-social Behaviour", Eltmean = "Employability")
lavaanPlot(model = SocSubMed, labels = labels4, node_options = list(shape = "box", fontname = "Helvetica"), edge_options = list(color = "grey"), coefs = TRUE, digits=3, stars="regress")

```

```{r}
#Bayes significance testing

Bftest1 <- BayesFactor::regressionBF(Polmean ~ ObjClass, rscaleCont = "medium", data = DATA_rel)
head(Bftest1, n=1)
Bftest2 <- BayesFactor::regressionBF(Eltmean ~ Polmean, rscaleCont = "medium", data = DATA_rel)
head(Bftest2, n=1)

Bftest3 <- BayesFactor::regressionBF(Socmean ~ ObjClass, rscaleCont = "medium", data = DATA_rel)
head(Bftest3, n=1)
Bftest4 <- BayesFactor::regressionBF(Eltmean ~ Socmean, rscaleCont = "medium", data = DATA_rel)
head(Bftest4, n=1)

Bftest5 <- BayesFactor::regressionBF(Polmean ~ SubClass, rscaleCont = "medium", data = DATA_rel)
head(Bftest5, n=1)
Bftest6 <- BayesFactor::regressionBF(Eltmean ~ Polmean, rscaleCont = "medium", data = DATA_rel)
head(Bftest6, n=1)

Bftest7 <- BayesFactor::regressionBF(Socmean ~ SubClass, rscaleCont = "medium", data = DATA_rel)
head(Bftest7, n=1)
Bftest8 <- BayesFactor::regressionBF(Eltmean ~ Socmean, rscaleCont = "medium", data = DATA_rel)
head(Bftest8, n=1)
```
